---
title: "Reproducible Research: Peer Assessment 1"
author: "angieFelipe"
date: "13/2/2018"
output: md_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



## Loading and preprocessing the data

First I will load the required packages
```{r packages, results='hide'}
paq<-c("dplyr","ggplot2","readr" )

lapply(paq, function(paq) {if (!require(paq, character.only=T)) 
{install.packages(paq);require(paq)}}) 
```




We unzip the data and read it. 

```{r get&clean}
zipF<- "activity.zip"
        unzip(zipF,overwrite = TRUE)

dataF<-"activity.csv"
        activity<-read_csv(dataF)
        
str(activity)                
```


## What is mean total number of steps taken per day?

In this part we will ignore the missing values in the dataset.

1. Make a histogram of the total number of steps taken each day

```{r hist_s_steps}

hist1<-activity %>% group_by(date) %>% summarise(s_steps=sum(steps,na.rm=T))
with(hist1, hist(s_steps, main = "Total Number of steps taken each day",
     xlab = "Number of steps by day"))

```




2. Calculate and report the **mean** and **median** total number of steps taken per day

```{r central_s_steps}
with_na <- c(mean=round(mean(hist1$s_steps),2), median = round(median(hist1$s_steps),2))
with_na

```




## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r intervalSteps}
plot1<-activity %>% group_by(interval) %>% summarise(a_steps=mean(steps,na.rm=T))
ggplot(plot1,aes(x=interval,y=a_steps) )+
        geom_line()+
        labs(title = "Average Number of steps by interval",
             y = "Average Number of steps")
```


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r maxIntervalSteps}
maxint <- plot1[plot1$a_steps == max(plot1$a_steps),1]
```

The interval is `r maxint`

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)

Ther are only missing values in the variable steps.

```{r naSteps}
activity$na <- is.na(activity$steps)
na <- sum(activity$na)
pct <- round(mean(activity$na)*100,2)
```

The total number of missing values in the dataset is `r na` which represents a
 `r pct` % of total observations.

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

For filling the missing values I will use the mean for that 5-minute interval in all days

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r noMissings}

plot1<-as.data.frame(plot1)
activity2<-as.data.frame(activity)
for (i in seq(1:nrow(activity2))){
        if (activity2[i,4]==TRUE){
                activity2[i,1]=plot1[plot1$interval==activity2[i,3],2]
        }
}

newna <- sum(is.na(activity2$steps))
newna
```

In the new dataset **activity2** total na are `r newna`.

4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. 
```{r daylysteps}
hist2<-activity2 %>% group_by(date) %>% summarise(s_steps=sum(steps,na.rm=T))
with(hist2, hist(s_steps, main = "Total Number of steps taken each day",
                 xlab = "Number of steps by day"))
without_na <- c(mean=round(mean(hist2$s_steps),2), median = round(median(hist2$s_steps),2))
without_na
```



*Do these values differ from the estimates from the first part of the assignment? *

```{r comparing_central}
table<-cbind(with_na,without_na)
table

```
 Both, mean and median have increase to the same value 
 
*What is the impact of imputing missing data on the estimates of the total daily number of steps?*

The distribution of steps by day is now more simetric and the mean and median are equal 

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r week}
activity2$wd<-weekdays(activity2$date)
wend <- unique(weekdays(activity2$date))[6:7]

for (i in seq(1:nrow(activity2))){
        if (activity2[i,5] %in% wend){
                activity2[i,5]="weekend"
        }
        else {
                activity2[i,5]="weekday"    
        }
}

activity2$wd <- as.factor(activity2$wd)

str(activity2$wd)
```

1. Make a panel plot containing a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r stepsbyweekday}
plot3 <- activity2 %>% group_by(wd,interval) %>%                         summarise(a_steps=mean(steps,na.rm=T))

ggplot(plot3, aes(x=interval, y=a_steps))+
        geom_line()+
        facet_grid(wd~.)+
        labs(y="Number of steps")
```

It looks as if there would be a different pattern in average steps by interval depending on weekdays or weekend.
